Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-30 latex/article.tex
Working on 4.

The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated \Jaco{triangulated} mesh. First note that
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangular
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ and thus
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

To implement the integration on $\Ball{R}{\vx} \cap \Shape$
(Fig.~\ref{fig:approx}-$a$), a naive approach consists in scanning all digital
points inside such domain (Fig.~\ref{fig:approx}-$b$) and then to estimate the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection. On GPU, \emph{mipmap} textures are used to
obtain a multi-resolution information: if the input binary object is stored in a
3D texture, GPU hardware can construct a multi-resolution pyramid such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
a given level $l$, a voxel $(x,y,z)$ contains the fraction of $\DSh$ belonging
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a lend of adaptive Gauss digitization process.

@@

Using the hierachical nature of the mipmap texture, we could also
consider hierarchical decompositions of either $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) or $\Ball{R}{\vx}\cap\Shape$
(Fig.~\ref{fig:approx2}-$b$). The idea would be to decompose the
regions into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. Note that the scenario in
Fig.~\ref{fig:approx2}-$b$, is clearly the best approach: the
volume is exact  (\wrt the regular grid case) since geometrical moments
are additive quantities, and the number of texture fetches is  optimal
at this point.

@@

\paragraph{Hierarchical Decomposition.}  As discussed above, best
results in terms of texture fetch can be obtained by the approach
presented in Fig.~\ref{fig:approx2}-$b$. At each vertex, the
computation algorithm would consist in a recursive decomposition of
the domain such that a cell is subdivided if it does not belong to
$\Ball{R}{\vx}\cap\Shape$. However, on the GPU %hardware,
this approach is
highly inefficient since the hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing the recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. Instead of implementing this approach,
we have considered a fast hierarchical approximation following
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierachical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan the ball cells and
estimate the area (or other moments) from the mipmap values associated
to this cell. Note that cells in the $B_R$ decomposition may not be
aligned with mipmap cells. However, we can use GPU %mipmap
functionalities which interpolates mipmap values if we probe at
non-discrete positions. For a given radius $R$, the expected number of
cells on $B_R$ is in $O(\log{R})$. Implementation details on the
hierarchical octree representation can be found in the supplementary
material (see Section \ref{sec:discussion}).

\paragraph{Dynamic Refinement.} In this approach, we want to optimize
the interactivity and the execution flow or parallelism on GPU. The
integration is simply computed using a regular grid at different
resolution $l$ (Fig. \ref{fig:approx}-$b$ and $c$). The shader code
becomes trivial (simple triple loop) and a lot of texture fetched are
involved, but interactivity can be easily obtained. Indeed, we
consider the following multi-pass approach:
\begin{enumerate}
\item When the surface geometry has been computed,we set the current
  level $l$ at an high level $l\EqDef l_{max}$ of the mipmap
  pyramid.
\item We compute the integrals (and the curvature tensor) at the
  mipmap level $l$ and  send the estimated quantities to the
  visualization stage.
\item If the user changes the camera position or the computation
  parameters, we return to step 1.
\item If the current framerate is above a given threshold, we decrease
  $l$ and return to step 2 if the final $l_0$ level has not been
  reached yet.
\end{enumerate}
A consequence of the multi-pass approach is that is there is no
interaction (camera settings, parameters), the GPU automatically
refine the estimation. Even if step 2 is quite expensive, in
$O\left(\left(\frac{R}{2^l}\right)^3\right)$, the interactive control
 in step 3 considerably improves the interactive exploration of the
 tensor with fast preliminary approximations which are quickly refined.

>>>>

The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated \Jaco{triangulated} mesh. First note that
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the \Jaco{triangulated}
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ \Jaco{for}
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

To implement the integration on $\Ball{R}{\vx} \cap \Shape$
(Fig.~\ref{fig:approx}-$a$), a naive approach consists in scanning all digital
points inside such domain (Fig.~\ref{fig:approx}-$b$) and then to estimate the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection. On GPU, \emph{mipmap} textures are used to
obtain a multi-resolution information: if the input binary object is stored in a
3D texture, GPU hardware can construct a multi-resolution pyramid such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
a given level $l$, a voxel $(x,y,z)$ contains the fraction of $\DSh$ belonging
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a \Jaco{blend ?} of adaptive Gauss digitization process.

@@

Using the hierachical nature of the mipmap texture, we could also
consider hierarchical decompositions of either $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) or $\Ball{R}{\vx}\cap\Shape$
(Fig.~\ref{fig:approx2}-$b$). The idea would be to decompose the
regions into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. Note that the scenario in
Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
volume is exact  (\wrt the regular grid case) since geometrical moments
are additive quantities, and the number of texture fetches is  optimal
at this point.

@@

\paragraph{Hierarchical Decomposition.}  As discussed above, best
results in terms of texture fetch can be obtained by the approach
presented in Fig.~\ref{fig:approx2}-$b$. At each vertex, the
computation algorithm would consist in a recursive decomposition of
the domain such that a cell is subdivided if it does not belong to
$\Ball{R}{\vx}\cap\Shape$. However, on the GPU %hardware,
this approach is
highly inefficient since the hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing the recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. Instead of implementing this approach,
we have considered a fast hierarchical approximation following
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierachical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan the ball cells and
estimate the area (or other moments) from the mipmap values associated
to this cell. Note that cells in the $B_R$ decomposition may not be
aligned with mipmap cells. However, we can use GPU %mipmap
functionalities which interpolates mipmap values if we probe at
non-discrete positions. For a given radius $R$, the expected number of
cells \Jaco{in} $B_R$ is in $O(\log{R})$. Implementation details on the
hierarchical octree representation can be found in the supplementary
material (see Section \ref{sec:discussion}).

\paragraph{Dynamic Refinement.} In this approach, we want to optimize
the interactivity and the execution flow or parallelism on GPU. The
integration is simply computed using a regular grid at different
resolution $l$ (Fig. \ref{fig:approx}-$b$ and $c$). The shader code
becomes trivial (simple triple loop) and a lot of texture \Jaco{fetches} are
involved, but interactivity can be easily obtained. Indeed, we
consider the following multi-pass approach:
\begin{enumerate}
\item When the surface geometry has been computed, we set the current
  level $l$ at an high level $l\EqDef l_{max}$ of the mipmap
  pyramid.
\item We compute the integrals (and the curvature tensor) at the
  mipmap level $l$ and  send the estimated quantities to the
  visualization stage.
\item If the user changes the camera position or the computation
  parameters, we return to step 1.
\item If the current framerate is above a given threshold, we decrease
  $l$ and return to step 2 if the final $l_0$ level has not been
  reached yet.
\end{enumerate}
A consequence of the multi-pass approach is that \Jaco{when} there is no
interaction (camera settings, parameters), the GPU automatically
refines the estimation. Even if step 2 is quite expensive, in
$O\left(\left(\frac{R}{2^l}\right)^3\right)$, the interactive control
 in step 3 considerably improves the interactive exploration of the
 tensor with fast preliminary approximations which are quickly refined.
